## Data Import and Preprocessing

### Bibliographic Data

All numbers in this section are counted by scripts, not typed. The scripts run in this order (see `README.md`):

1. `R/prisma_pipeline.R` reads the Zotero exports (`bib/bib_raw/*.bib`), removes duplicates and applies the manual and automatic screening.
2. `R/enrich_records.R` fills missing abstracts and DOIs (OpenAlex, Crossref) and checks every DOI against Crossref.
3. `R/review_weak_matches.R` applies rules to the weak DOI matches.
4. `R/prisma_counts_final.R` writes the final counts that are shown here.

```{r load_prisma, echo=FALSE}
# Read the outputs of the pipeline. Nothing is recomputed in this report.
prisma_files <- c("data/prisma/prisma_counts_full.csv",
                  "data/prisma/records_enriched.csv",
                  "data/prisma/records_full.rds")

# stop with a clear message if the pipeline has not been run
missing_files <- prisma_files[!file.exists(prisma_files)]
if (length(missing_files) > 0) {
  stop("Run the pipeline first (see README.md). Missing: ", paste(missing_files, collapse = ", "))
}

prisma   <- readr::read_csv("data/prisma/prisma_counts_full.csv", show_col_types = FALSE)
enriched <- readr::read_csv("data/prisma/records_enriched.csv",   show_col_types = FALSE)
full     <- readRDS("data/prisma/records_full.rds")

# helpers: n of the row whose item starts with a text; number with thousand marks
pn <- function(start) prisma$n[startsWith(prisma$item, start)][1]
fm <- function(x) format(x, big.mark = ",")

# some abstracts hold control characters / Unicode non-characters that break tolower();
# replace them with a space
clean_txt <- function(x) stringi::stri_replace_all_regex(x, "[\\p{C}]", " ")

# one table of included records, with the column names the analysis below expects
bib_data <- enriched %>%
  dplyr::left_join(dplyr::select(full, id, keywords), by = "id") %>%
  dplyr::transmute(
    BIBTEXKEY = key,
    TITLE     = clean_txt(title),
    AUTHOR    = clean_txt(author),
    YEAR      = as.numeric(year),
    JOURNAL   = journal,
    DOI       = doi_final,
    ABSTRACT  = clean_txt(abstract_final),
    KEYWORDS  = clean_txt(keywords),
    CROSSREF  = crossref_status
  )
```

Our Zotero export holds `r fm(pn("Records in Zotero"))` records. After duplicate removal and screening, `r fm(pn("TOTAL included"))` records are included.

```{r prisma_flow, echo=FALSE, fig.width=9, fig.height=7, fig.cap="PRISMA flow, counted by R/prisma_pipeline.R"}
source("R/plot_prisma.R")
plot_prisma(prisma)
```

#### Duplicates

We removed `r fm(sum(prisma$n[prisma$stage == "dedup"]))` records in four steps. We keep one copy of each work and remove only the extra copies.

- Same Zotero citekey, exported twice (mostly the same records in `documents.bib` and `reports.bib`): `r fm(pn("Removed: same citekey"))`.
- Same DOI: `r pn("Removed: same DOI")`.
- Same title and year: `r pn("Removed: same title")`.
- Same work with a different year or version (preprint and published article, wrong year, garbled author), checked by hand: `r pn("Removed: same work")` (`data/prisma/duplicates_reviewed.csv`).

#### Screening

We removed `r pn("Removed: manual")` records marked `bin` in Zotero, and `r sum(prisma$n[startsWith(prisma$item, "Removed: auto")])` records by automatic rules (annual reports, syllabi, catalogues, non-scientific sources). Only `r pn("Included (manual")` of the included records were marked `read` by hand; the rest have not been screened for relevance yet.

#### Crossref check, abstracts and DOIs

We checked every DOI against Crossref (title and year must match). `r fm(pn("Verified"))` of the `r fm(pn("Included records checked"))` included records are verified. `r fm(pn("Not verified: no DOI"))` have no DOI, so they cannot be checked.

Abstracts: `r fm(pn("Abstract: total"))` of the included records have one (`r fm(pn("Abstract: already"))` came from Zotero, `r fm(pn("Abstract: added"))` were added from OpenAlex and Crossref).

```{r prisma_table, echo=FALSE}
# the full count table
prisma %>%
  dplyr::select(Stage = stage, Item = item, n) %>%
  dplyr::mutate(n = format(n, big.mark = ",")) %>%
  knitr::kable(caption = "PRISMA counts (data/prisma/prisma_counts_full.csv)") %>%
  kableExtra::kable_styling(bootstrap_options = c("striped", "hover"), font_size = 12)
```
