## Data Import and Preprocessing

### Bibliographic Data

Import bibliographic records using the `bib2df`, `qs` and `qs2` packages [@R-bib2df, @R-qs, @R-qs2], converting our BibTeX file into a structured data frame for systematic processing.

```{r read_bib_data, echo=FALSE}

# Work on Zotero in Lintner (prepare and pre-process)
# Export properly from Zotero using Better BibTeX plugin
# Or configure Zotero's BibTeX export settings

# Stream processing for large files
clean_bib_files <- function(input_dir, output_file) {
  bib_files <- list.files(input_dir, pattern = "\\.bib$", full.names = TRUE)
  
  # Process files sequentially, not all in memory
  processed_entries <- character(0)
  
  for (bib_file in bib_files) {
    message("Processing: ", basename(bib_file))
    
    tryCatch({
      # Use proper BibTeX parser
      parsed <- bibtex::read.bib(bib_file)
      
      if (length(parsed) > 0) {
        # Write valid entries immediately
        bibtex::write.bib(parsed, file = output_file, append = TRUE)
      }
    }, error = function(e) {
      warning("Failed to process ", bib_file, ": ", e$message)
    })
  }
}
```

Although `bibtex` `read.bib` and `write.bib` functions [@R-bibtex] have done much of the work of removing irrelevant work, we still do some post-processisng to esure that we have a meaningful collection of papers to proceed. We combine all qs bib files into a single new bib file.

```{r create_qs_collection, echo=FALSE}
# Use data structures from the start
# run fresh populate bib/bib_raw/ with zotero exports
# delete "bib/Bib_clean.bib" and "qs/Bib_valid_only.qs"
if (!file.exists("qs/Bib_valid_only.qs")) {
  clean_bib_files("bib/bib_raw/", "bib/Bib_clean.bib")
  bib_data <- bib2df::bib2df("bib/Bib_clean.bib")
  qs::qsave(bib_data, "qs/Bib_valid_only.qs")
} else {
  bib_data <- qs::qread("qs/Bib_valid_only.qs")
}
```

Organize all bib data into an accessible dataset with `dplyr` and the `purrr` function `map_dfr` [@R-dplyr, @R-purrr].

```{r combined_bib_data_qs}

  # Standardize column types to prevent binding errors
  combined_bib_data <- bib_data %>%
    dplyr::mutate(
      # Convert YEAR to character first, then to numeric (handles NAs)
      YEAR = as.character(YEAR),
      YEAR = as.numeric(YEAR),
      # Ensure other common columns are consistent
      TITLE = as.character(TITLE),
      ANNOTE = as.character(ANNOTE),
      BIBTEXKEY = as.character(BIBTEXKEY)
    )
```

Our BibTeX contains `r nrow(combined_bib_data)` citations and `r ncol(combined_bib_data)` reference information items for each.

### BibTeX Cleaning

We prepare the BibTeX file using `dplyr` [@R-dplyr], for annotation tracking and text consolidation.

```{r clean_bib_data, echo=FALSE}
# Basic cleaning and text preparation
clean_bib_data <- combined_bib_data %>%
  dplyr::mutate(
    has_annotation = !is.na(ANNOTE) & ANNOTE != "",
    annotation_text = ifelse(has_annotation, ANNOTE, ""),
    full_text = paste(TITLE, ANNOTE, sep = " "),
    year = as.numeric(YEAR)
  )
```

`r length(clean_bib_data$full_text)` of the citations have text to work with.

## Run PRISMA Pipeline

Detect duplicates with `identify_duplicates`, using `dplyr` [@R-dplyr] and base R duplicate detection [@R-base] to find both title-based and DOI-based duplicates through text normalization and exact matching.

```{r identify_duplicates, echo=FALSE}
identify_duplicates <- function(bib_data) {
  bib_data %>%
    mutate(
      # Simple title-based duplicate detection
      title_clean = str_remove_all(tolower(TITLE), "[^a-z0-9 ]"),
      is_duplicate = duplicated(title_clean) | duplicated(title_clean, fromLast = TRUE),
      
      # DOI-based duplicate detection
      is_doi_duplicate = !is.na(DOI) & (duplicated(DOI) | duplicated(DOI, fromLast = TRUE)),
      
      # Combined duplicate flag
      duplicate_status = case_when(
        is_doi_duplicate ~ "doi_duplicate",
        is_duplicate ~ "title_duplicate", 
        TRUE ~ "unique"
      )
    )
}

 # Duplicate identification and REMOVAL
  duplicates_summary <- identify_duplicates(clean_bib_data)
  
  # Remove duplicates
  duplicates_removed <- duplicates_summary %>% filter(duplicate_status == "unique")
```

`r nrow(duplicates_removed)` duplicates were removed. 

Apply manual classifications from bib keywords using `apply_manual_classifications`, with `dplyr` [@R-dplyr] and `stringr` [@R-stringr] to identify papers that we manually marked 'bin' for exclusion or 'read' for inclusion.

```{r apply_manual_classifications, echo=FALSE}
apply_manual_classifications <- function(bib_data) {
  bib_data %>%
    mutate(
      has_bin = str_detect(tolower(KEYWORDS), "\\bbin\\b"),
      has_read = str_detect(tolower(KEYWORDS), "read cw|read pka"),
      
      manual_decision = case_when(
        has_bin ~ "exclude",
        has_read & !has_bin ~ "include",
        TRUE ~ "unreviewed"
      ),
      
      manual_reason = case_when(
        has_bin ~ "curator_marked_bin",
        has_read ~ "manually_reviewed_keep",
        TRUE ~ "not_reviewed"
      )
    )
}

# Manual classifications and REMOVAL
  manual_classification_summary <- apply_manual_classifications(duplicates_removed)
  # Remove manually excluded records
  removed_manually <- manual_classification_summary %>% filter(manual_decision != "exclude")

```

`r nrow(removed_manually)` records were removed manually. 

Apply additional exclusion rules. Remove all annual reports, syllabus, magazines, course notes, legal documents, bibliography collections etc.

```{r apply_exclusion_rules, echo=FALSE}
apply_exclusion_rules <- function(bib_data) {
  bib_data %>%
    mutate( 
      # Document type exclusions (specific patterns)
      exclude_doc_type = str_detect(tolower(TITLE), 
                                    "annual report|syllabus|catalog|course notes|legal document|bibliography collection"),
      # Scientific content check (be more specific)
      exclude_non_scientific = case_when(
        !is.na(JOURNAL) & str_detect(tolower(JOURNAL), 
                                     "legal review|law journal|news|magazine|newspaper") ~ TRUE,
        str_detect(tolower(TITLE), "^news:|^magazine:|^blog:|newspaper article") ~ TRUE,
        TRUE ~ FALSE
      ),
      # Preprint handling - only exclude if we're confident it's a duplicate
      exclude_preprint = case_when(
        !is.na(JOURNAL) & str_detect(tolower(JOURNAL), "arxiv|preprint|biorxiv|medrxiv") & 
          !is.na(PUBLISHER) & str_detect(tolower(PUBLISHER), "elsevier|springer|wiley|taylor") ~ TRUE,
        TRUE ~ FALSE
      ),
      # Content relevance - be more inclusive, focus on clear exclusions
      exclude_content = case_when(
        # Only exclude if clearly not decision-related
        str_detect(tolower(TITLE), "chemistry experiment|physics lab|pure mathematics|organic synthesis") ~ TRUE,
        # Keep anything that might be decision-related
        str_detect(tolower(TITLE), "decision|choice|select|evaluate|assess|analyze|model|uncertain|risk") ~ FALSE,
        # Default to include for systematic review
        TRUE ~ FALSE
      ),
      # Combined automated exclusion
      automated_exclude = exclude_doc_type | #exclude_language | 
        exclude_non_scientific | exclude_preprint | exclude_content,
      automated_reason = case_when(
        exclude_doc_type ~ "invalid_document_type",
       # exclude_language ~ "non_english",
        exclude_non_scientific ~ "non_scientific",
        exclude_preprint ~ "preprint_with_published_version", 
        exclude_content ~ "irrelevant_content",
        TRUE ~ "not_excluded"
      )
    )
}


  # Automated exclusion rules and REMOVAL
  excluded_bib <- apply_exclusion_rules(removed_manually)
  
  # Remove automated exclusions (only from unreviewed records)
  excluded_irrelevant <- excluded_bib %>% 
    filter(!(manual_decision == "unreviewed" & automated_exclude))
```

Additional exclusion rules removed `r nrow(excluded_irrelevant)` records.

Create `clean_bib_data_prisma` with all classified data. 

```{r clean_bib_data_prisma_excluded_irrelevant}
clean_bib_data_prisma <- excluded_irrelevant
```
