---
title: "Interactive R Analysis Example - Hosyana Review"
output:
  html_document:
    code_folding: show
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

# Interactive R Analysis Example

This R Markdown document demonstrates interactive line-by-line analysis of the comprehensive bibliography data. In RStudio, you can run each chunk individually by clicking the green arrow or pressing `Ctrl+Shift+Enter`, just like your usual workflow.

**Usage in RStudio:**
- Click on a line and press `Ctrl+Enter` to run that line
- Run entire chunks with `Ctrl+Shift+Enter`
- Variables persist between chunks
- Results display immediately in the console or inline

## Setup and Data Loading

```{r libraries}
# Load required libraries (run this chunk first)
library(dplyr)
library(ggplot2)
library(readr)
library(tidyr)
library(stringr)
library(DT)
library(plotly)
library(knitr)
library(kableExtra)

cat("✓ Libraries loaded successfully\n")
```

```{r load_data}
# Load the comprehensive data
df <- read_csv('FINAL_methods_analysis.csv', show_col_types = FALSE)
cat("Loaded", nrow(df), "papers\n")
cat("Columns:", paste(names(df)[1:10], collapse = ", "), "...\n")
```

```{r data_preview}
# Quick data overview
head(df, 3) %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover"))
```

## Interactive Data Exploration

```{r data_quality}
# Check data quality - run this line to see missing data
missing_data <- df %>%
  summarise_all(~sum(is.na(.))) %>%
  pivot_longer(everything(), names_to = "column", values_to = "missing") %>%
  filter(missing > 0) %>%
  arrange(desc(missing))

if(nrow(missing_data) > 0) {
  cat("Missing data by column:\n")
  print(missing_data)
} else {
  cat("No missing data found!\n")
}
```

```{r year_distribution}
# Year distribution - run this to see publication years
year_counts <- df %>% count(YEAR, sort = TRUE)
cat("Year range:", min(df$YEAR, na.rm = TRUE), "-", max(df$YEAR, na.rm = TRUE), "\n")
cat("Most productive year:", year_counts$YEAR[1], "with", year_counts$n[1], "papers\n")
```

```{r method_detection}
# Method detection summary - run this line by line to explore
papers_with_methods <- df %>% filter(HAS_METHODS == 1)
detection_rate <- nrow(papers_with_methods) / nrow(df) * 100

cat("Papers with methods detected:", nrow(papers_with_methods), "\n")
cat("Detection rate:", round(detection_rate, 1), "%\n")
```

## Method Analysis (Interactive)

```{r method_columns}
# Get method columns
method_columns <- c(
  'DECISION_ANALYSIS', 'POLICY_INTERVENTION', 'UNCERTAINTY_ANALYSIS',
  'STAKEHOLDER_EXPERT', 'MODELING_SIMULATION', 'BAYESIAN_PROBABILISTIC',
  'COMPUTER_ASSISTED', 'VALUE_INFORMATION', 'MULTI_CRITERIA', 'OPTIMIZATION',
  'ECONOMIC_EVALUATION', 'GAME_THEORY', 'BEHAVIORAL_PSYCHOLOGY',
  'SYSTEMS_COMPLEXITY', 'TECHNOLOGY_INNOVATION', 'ENVIRONMENTAL_SUSTAINABILITY',
  'QUALITY_PERFORMANCE', 'RISK_SAFETY', 'FORECASTING_PREDICTION', 'EVALUATION_ASSESSMENT'
)

cat("Found", length(method_columns), "method categories\n")
```

```{r method_popularity}
# Calculate method popularity (run this to see top methods)
method_counts <- df %>%
  select(all_of(method_columns)) %>%
  summarise_all(sum) %>%
  pivot_longer(everything(), names_to = "method", values_to = "count") %>%
  arrange(desc(count)) %>%
  mutate(clean_name = str_replace_all(method, "_", " ") %>% str_to_title())

cat("Top 10 methods:\n")
for(i in 1:10) {
  cat("  ", method_counts$clean_name[i], ":", method_counts$count[i], "papers\n")
}
```

```{r method_visualization}
# Quick visualization of top methods
top_methods_plot <- method_counts %>%
  head(8) %>%
  ggplot(aes(x = reorder(clean_name, count), y = count)) +
  geom_col(fill = "steelblue", alpha = 0.7) +
  coord_flip() +
  labs(title = "Top 8 Decision Support Methods",
       x = "Method",
       y = "Number of Papers") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))

print(top_methods_plot)
```

## Temporal Analysis (Decade Trends)

```{r decade_analysis}
# Create decade column
df <- df %>%
  mutate(
    decade = floor(YEAR / 10) * 10,
    decade = ifelse(is.na(decade), 2020, decade)  # Handle missing years
  )

# Look at decade distribution
decade_counts <- df %>%
  count(decade, sort = FALSE) %>%
  filter(!is.na(decade))

cat("Papers by decade:\n")
for(i in 1:nrow(decade_counts)) {
  cat("  ", decade_counts$decade[i], "s:", decade_counts$n[i], "papers\n")
}
```

```{r specific_decade}
# Focus on specific decade (change the decade here)
decade_to_analyze <- 2010
decade_data <- df %>% filter(decade == decade_to_analyze)

cat("\n", decade_to_analyze, "s Analysis:\n", sep = "")
cat("Total papers:", nrow(decade_data), "\n")
cat("Papers with methods:", sum(decade_data$HAS_METHODS == 1), "\n")
```

```{r decade_methods}
# Top methods for this decade
decade_methods <- decade_data %>%
  select(all_of(method_columns)) %>%
  summarise_all(sum) %>%
  pivot_longer(everything(), names_to = "method", values_to = "count") %>%
  arrange(desc(count)) %>%
  mutate(clean_name = str_replace_all(method, "_", " ") %>% str_to_title())

cat("\nTop 5 methods in", decade_to_analyze, "s:\n")
for(i in 1:5) {
  cat("  ", decade_methods$clean_name[i], ":", decade_methods$count[i], "papers\n")
}
```

## Recent Trends (2020-2024)

```{r recent_analysis}
# Focus on recent years (2020-2024)
recent_data <- df %>% filter(YEAR >= 2020 & YEAR <= 2024)
cat("Recent papers (2020-2024):", nrow(recent_data), "\n")

# Year by year breakdown
recent_by_year <- recent_data %>%
  count(YEAR, sort = FALSE)

cat("\nYear-by-year breakdown:\n")
for(i in 1:nrow(recent_by_year)) {
  cat("  ", recent_by_year$YEAR[i], ":", recent_by_year$n[i], "papers\n")
}
```

```{r recent_methods}
# Top methods in recent period
recent_methods <- recent_data %>%
  select(all_of(method_columns)) %>%
  summarise_all(sum) %>%
  pivot_longer(everything(), names_to = "method", values_to = "count") %>%
  arrange(desc(count)) %>%
  mutate(
    clean_name = str_replace_all(method, "_", " ") %>% str_to_title(),
    percentage = count / sum(recent_data$HAS_METHODS == 1) * 100
  )

cat("Top methods in 2020-2024:\n")
for(i in 1:8) {
  cat("  ", recent_methods$clean_name[i], ":", recent_methods$count[i],
      "papers (", round(recent_methods$percentage[i], 1), "%)\n")
}
```

## Custom Analysis (Edit and Run)

```{r multi_method}
# Example: Find papers using multiple methods
df <- df %>%
  rowwise() %>%
  mutate(method_count = sum(c_across(all_of(method_columns)))) %>%
  ungroup()

multi_method <- df %>% filter(method_count > 2)

cat("Papers using 3+ methods:", nrow(multi_method), "\n")
cat("Average methods per paper:", round(mean(df$method_count), 2), "\n")
```

```{r search_titles}
# Example: Search for specific terms in titles
search_term <- "decision"  # Change this to search for different terms
matching_papers <- df %>%
  filter(str_detect(tolower(TITLE), tolower(search_term)))

cat("Papers with '", search_term, "' in title:", nrow(matching_papers), "\n")
if(nrow(matching_papers) > 0) {
  cat("\nExample titles:\n")
  for(i in 1:min(3, nrow(matching_papers))) {
    cat("  -", matching_papers$TITLE[i], "\n")
  }
}
```

```{r trend_plot}
# Example: Create a simple trend plot
method_to_track <- 'DECISION_ANALYSIS'  # Change this to track different methods

trend_data <- df %>%
  filter(decade >= 1970) %>%
  group_by(decade) %>%
  summarise(method_count = sum(.data[[method_to_track]]), .groups = "drop")

trend_plot <- trend_data %>%
  ggplot(aes(x = decade, y = method_count)) +
  geom_line(linewidth = 1.2, color = "steelblue") +
  geom_point(size = 3, color = "darkblue") +
  labs(
    title = paste("Trend:", str_replace_all(method_to_track, "_", " ") %>% str_to_title()),
    x = "Decade",
    y = "Number of Papers"
  ) +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))

print(trend_plot)
```

## Interactive Data Table

```{r interactive_table}
# Create an interactive table of recent papers
recent_summary <- recent_data %>%
  filter(HAS_METHODS == 1) %>%
  select(TITLE, YEAR, AUTHORS, JOURNAL, CONFIDENCE) %>%
  arrange(desc(YEAR), desc(CONFIDENCE)) %>%
  head(50)

# Create interactive table
datatable(recent_summary,
          options = list(pageLength = 10, scrollX = TRUE),
          caption = "Recent Papers with Detected Methods (2020-2024)")
```

## Summary Statistics

```{r summary_stats}
# Generate summary statistics
summary_stats <- list(
  total_papers = nrow(df),
  papers_with_methods = sum(df$HAS_METHODS == 1),
  detection_rate = round(sum(df$HAS_METHODS == 1) / nrow(df) * 100, 1),
  year_range = paste(min(df$YEAR, na.rm = TRUE), "-", max(df$YEAR, na.rm = TRUE)),
  top_method = method_counts$clean_name[1],
  top_method_count = method_counts$count[1],
  multi_method_papers = sum(df$method_count > 1),
  avg_methods_per_paper = round(mean(df$method_count), 2)
)

cat("=== SUMMARY STATISTICS ===\n")
cat("Total papers analyzed:", summary_stats$total_papers, "\n")
cat("Papers with methods detected:", summary_stats$papers_with_methods, "\n")
cat("Detection rate:", summary_stats$detection_rate, "%\n")
cat("Year coverage:", summary_stats$year_range, "\n")
cat("Top method:", summary_stats$top_method, "(", summary_stats$top_method_count, "papers)\n")
cat("Multi-method papers:", summary_stats$multi_method_papers, "\n")
cat("Average methods per paper:", summary_stats$avg_methods_per_paper, "\n")
```

---

## How to Use This Interactive Analysis

**In RStudio:**

1. **Run individual lines** with `Ctrl+Enter` to see immediate results in the console
2. **Run entire chunks** with `Ctrl+Shift+Enter` or click the green arrow
3. **Modify variables** in chunks above (like `decade_to_analyze`, `search_term`, `method_to_track`) and re-run
4. **Variables persist** between chunks - you can reference `df`, `method_counts`, etc. in new chunks
5. **Create new chunks** with `Ctrl+Alt+I` to experiment with your own analysis

**Key Interactive Features:**
- All variables remain in your R environment for further analysis
- Plots appear in the Plots pane
- Data tables show in the Viewer pane
- You can modify any parameters and instantly re-run to see changes

This gives you the exact same interactive experience as your usual RStudio workflow!
